---
title: "R Coding Exercise"
---

load the dslabs and ggplot2 library

```{r}
library(dslabs)
library(ggplot2)
```

help for gapminder and overview of data structure

```{r}
help(gapminder)

str(gapminder)
```

Summary of the gapminder data

```{r}
summary(gapminder)
```

What type of object is gapminder

```{r}
class(gapminder)
```

Filter out the data to include only African Countries and check the structure of the new data set.

```{r}
africadata <- gapminder[gapminder$continent == "Africa", ]
str(africadata)
summary(africadata)
```

creating 2 new objects from the africadata data set

```{r}
africa_infant_life <- africadata[, c("infant_mortality", "life_expectancy")]
africa_pop_life <- africadata[, c("population", "life_expectancy")]
```

Scatter plot for life expectancy as a function of infant mortality

```{r}
ggplot(africa_infant_life, aes(x = infant_mortality, y = life_expectancy)) +
  geom_point() +
  labs(title = "Life Expectancy vs. Infant Mortality in African Countries",
       x = "Infant Mortality",
       y = "Life Expectancy") 
```

Scatter plot for life expectancy as a function of population size

```{r}
ggplot(africa_pop_life, aes(x = population, y = life_expectancy)) +
  geom_point() +
  scale_x_log10() +
  labs(title = "Life Expectancy vs. Population Size in African Countries",
       x = "Population (log scale)",
       y = "Life Expectancy")
```

referring back to the African data set to observe what is happening.

```{r}
str(africadata)
```

The streaks of data can be a result of there being different years for the same countries, which could be a reason why you see streaks in different areas of the map, as it could be plotting the same country year by year.

Check for missing values in infant_mortality by year

```{r}
missing_infant_mortality <- aggregate(infant_mortality ~ year, data = africadata, function(x) sum(is.na(x)))
print(missing_infant_mortality)
```

filter for the year 2000 and display the structure and summary of the data

```{r}
africa_2000 <- africadata[africadata$year == 2000, ]

str(africa_2000)

summary(africa_2000)
```

Scatter plot of life expectancy vs. infant mortality for the year 2000

```{r}
ggplot(africa_2000, aes(x = infant_mortality, y = life_expectancy)) +
  geom_point() +
  labs(title = "Life Expectancy vs. Infant Mortality in African Countries (2000)",
       x = "Infant Mortality",
       y = "Life Expectancy")
```

Scatter plot of life expectancy vs. population size for the year 2000

```{r}
ggplot(africa_2000, aes(x = population, y = life_expectancy)) +
  geom_point() +
  scale_x_log10() +
  labs(title = "Life Expectancy vs. Population Size in African Countries (2000)",
       x = "Population (log scale)",
       y = "Life Expectancy")
```

Fit linear models

```{r}
fit1 <- lm(life_expectancy ~ infant_mortality, data = africa_2000)
fit2 <- lm(life_expectancy ~ population, data = africa_2000)
```

summary of the fitted models

```{r}
summary(fit1)
summary(fit2)
```

Given that the p-value for the first fitted model (fit1) is extremely low, we can conclude that there is a significant negative relationship between infant mortality and life expectancy.

Given the p-value for the second fitted model (fit2) is high, we can conclude that there is no significant relationship between population size and life expectancy.

This section is from My other account :)

Load the death_prob dataset, which is a dataset that lists the probability of death within 1 year by age and sex in the United States in 2015.
Also get an overview of the dataset structure
```{r}
data(death_prob)
class(death_prob)
str(death_prob)
```

Next I will only be looking at the death probability of just women, so I will create a new dataset with only the "Feamale" variable.
I also repeat the steps from death_pr to look at the new dataset structure
```{r}
women_death_prob <- subset(death_prob, sex == "Female")
str(women_death_prob)
summary(women_death_prob)
```

Create new datasets from the women_depth_prob object.
```{r}
age_death_prob <- women_death_prob[, c("age", "death_prob")]
year_death_prob <- women_death_prob[, c("year", "death_prob")]
```

display the new objects' structure
```{r}
str(age_death_prob)
str(year_death_prob)
```
Display the summary of both objects'
```{r}
summary(age_death_prob)
summary(year_death_prob)
```

Plot death_prob as a function of age
```{r}
ggplot(age_death_prob, aes(x = age, y = death_prob)) +
  geom_point() +
  labs(title = "Death Probability as a Function of Age",
       x = "Age",
       y = "Death Probability")
```
Plot death_prob as a function of year
```{r}
ggplot(year_death_prob, aes(x = year, y = death_prob)) +
  geom_point() +
  labs(title = "Death Probability as a Function of Year",
       x = "Year",
       y = "Death Probability")
```
Fit linear model: death_prob as a function of age and the second model as a function of the year
```{r}
fit_age <- lm(death_prob ~ age, data = age_death_prob)
fit_year <- lm(death_prob ~ year, data = year_death_prob)
```
Here is a Summary of the 2 new models
```{r}
summary(fit_age)
summary(fit_year)
```
