---
title: "Data Exercise"
---

###Option 2

#Generate the Sythetic Data
For this I will do Age, salary, and employee_type for my variables
```{r}

library(tidyverse)
library(ggplot2)

set.seed(123)
n <- 1000

# Continuous variables
x1 <- rnorm(n, mean = 50, sd = 10)  
x2 <- rnorm(n, mean = 30, sd = 5)   

# Categorical variable with 3 levels
x3 <- sample(c("A", "B", "C"), n, replace = TRUE)

# Dependent variable with some associations
y <- 3 + 2*x1 + 0.5*x2 + ifelse(x3 == "A", 5, ifelse(x3 == "B", -3, 0)) + rnorm(n)

data <- data.frame(x1, x2, x3, y)
head(data)
```


#Explore the data
```{r}
summary(data)
```
```{r}
ggplot(data, aes(y)) +
  geom_histogram(binwidth = 10, fill = 'blue', color = 'black', alpha = 0.7) +
  ggtitle('Histogram of y')

ggplot(data, aes(x = x1, y = y)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "red") +
  ggtitle('Scatter plot of y vs x1')

ggplot(data, aes(x = x2, y = y)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "red") +
  ggtitle('Scatter plot of y vs x2')

ggplot(data, aes(x = x3, y = y)) +
  geom_boxplot(fill = 'cyan', color = 'black') +
  ggtitle('Boxplot of y vs x3')
```


#Simple models for generated data

```{r}
#Linear Models
model1 <- lm(y ~ x1 + x2 + x3, data = data)

model2 <- lm(y ~ x1 * x3 + x2, data = data)

# Logistic regression (create a binary outcome variable)
data$y_bin <- ifelse(data$y > median(data$y), 1, 0)
model3 <- glm(y_bin ~ x1 + x2 + x3, data = data, family = binomial)

# Print model summaries to compare
summary(model1)
summary(model2)
summary(model3)
```
Model 1 shows the main linear associations, while model 2 shows both main and interaction effects.
Model 3 demonstrates the flexibility of logistic regression in handling binary outcomes with the same associations in the previous 2 models.





